Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-06-02 neurips-20/cvx-flats.tex
Pass through sec 4

\botodo{I felt the reader needed a bit more transition here, and added one. Also I think this proof sketch was much too dense for all but the most expert in convex analysis! I slowed things down, and see next note.}
Thanks to Theorem \ref{thm:consistent-implies-indir-elic}, the task is now this: given a target property $\gamma$, or loss $\ell$ with $\gamma = \prop{\ell}$, prove lower bounds on any $d$ such that some convex $L: \reals^d \times \Y \to \reals$ indirectly elicits $\gamma$.
We now explore a tool, Theorem \ref{thm:cvx-flats}, for proving such convex-elicitation lower bounds.
The key idea, crystallized from the proof of \citet[Theorem 16]{ramaswamy2016convex}, is to consider a particular optimal surrogate report $u \in \reals^d$.
\botodo{wording/credit ok?}
Theorem \ref{thm:cvx-flats} will show that if $d$ is small, then the level set $\{p : u \in \argmin_{u'} L(u',p)\}$ \bo{used comma instead of semicolon} must be large; in fact, it must roughly contain a high-dimensional affine subspace, or \emph{flat}.
But by definition of indirect elicitation, there is some level set $\gamma_r$ (where $u$ is linked to $r$) containing this flat as well.
The point of this result is the contrapositive: if $\gamma$ has a more intricate level set that does not contain any high-dimensional flats, then $\gamma$ cannot have a low-dimensional consistent surrogate.

\botodo{Not sure the proof sketch is worth it, since the proof itself is pretty short. Commenting out for now, but let me know what y'all would prefer!}
\botodo{TODO: we have not yet introduced subgradients or the subdifferential!}
%As a first step, observe that all minima of convex functions are global, so $\vec 0  \in \partial L(u, y)$ if and only if $u$ minimizes $L$ in its first component.
%This allows us to consider subgradient sets of the loss at a fixed distribution $p$ as the weighted Minkowski sums of subgradient sets for the loss on each outcome.\bo{This won't be clear except to readers very well versed in convex analysis. I think we'll have to sketch why that's true.}
%Fixing a distribution $p \in \simplex$, we construct a set of $n$ witnesses in $\partial L(u,y)$ for each $y$ whose $p$-weighted sum is $\vec 0$.
%The nullspace of the matrix of these witnesses yields a flat, formally introduced below, of codimension at most $d$ intersecting our fixed distribution $p$ and the level set(s) $\gamma_r$ containing $p$.

@@

\bo{I like all this, but maybe it can be reworded to flow better with the above?}
We now consider subgradients of convex losses and their flats in order to understand convex elicitation complexity. 
The intuition for this result comes from an observation about the existence of a set of witnesses in the Minkowski sum from the proof of~\citet[Theorem 16]{ramaswamy2016convex}, although indirect elicitation lets us simplify their proof that originally relied on calibration.
In fact, their main proof tool, the subspace of feasible dimensions (Definition~\ref{def:subspace-feas}), is a linear shift of a flat satisfying the requirements of Theorem~\ref{thm:cvx-flats}, which lets us deduce their result as Corollary~\ref{cor:fsd-bound}.

@@

%Since Theorem~\ref{thm:consistent-implies-indir-elic} says that consistency implies indirect elicitation, we now apply results on convex elicitation complexity, like the result above, to understand the minimal dimension required to construct a consistent surrogate loss.
%This result suggests that elicitation complexity greater than $d$ implies no indirect elicitation of the property of interest (or target loss eliciting such a property), which in turn implies there is no $d$-dimensional consistent surrogate with respect to a given loss or property of interest.
\bo{I felt this remark was redundant now, so I removed it.}

>>>>

\botodo{I felt the reader needed a bit more transition here, and added one. Also I think this proof sketch was much too dense for all but the most expert in convex analysis! I slowed things down, and see next note.}
Thanks to Theorem \ref{thm:consistent-implies-indir-elic}, the task is now this: given a target property $\gamma$, or loss $\ell$ with $\gamma = \prop{\ell}$, prove lower bounds on any $d$ such that some convex $L: \reals^d \times \Y \to \reals$ indirectly elicits $\gamma$.
We now explore a tool, Theorem \ref{thm:cvx-flats}, for proving such convex-elicitation lower bounds.
The key idea, crystallized from the proof of \citet[Theorem 16]{ramaswamy2016convex}, is to consider a particular optimal surrogate report $u \in \reals^d$.
\botodo{wording/credit ok? \jessie{I think so}}
Theorem \ref{thm:cvx-flats} will show that if $d$ is small, then the level set $\{p : u \in \argmin_{u'} \exploss{L}{u'}{p}\}$ \bo{used comma instead of semicolon} must be large; in fact, it must roughly contain a high-dimensional affine subspace, or \emph{flat}.
But by definition of indirect elicitation, there is some level set $\gamma_r$ (where $u$ is linked to $r$) containing this flat as well.
The point of this result is the contrapositive: if $\gamma$ has a more intricate level set that does not contain any high-dimensional flats, then $\gamma$ cannot have a low-dimensional consistent surrogate.

\botodo{Not sure the proof sketch is worth it, since the proof itself is pretty short. Commenting out for now, but let me know what y'all would prefer!}
\botodo{TODO: we have not yet introduced subgradients or the subdifferential!}\jessiet{Added as a footnote since I would hope most people know what a subgradient is, but also a good point to not assume.}
%As a first step, observe that all minima of convex functions are global, so $\vec 0  \in \partial L(u, y)$ if and only if $u$ minimizes $L$ in its first component.
%This allows us to consider subgradient sets of the loss at a fixed distribution $p$ as the weighted Minkowski sums of subgradient sets for the loss on each outcome.\bo{This won't be clear except to readers very well versed in convex analysis. I think we'll have to sketch why that's true.}
%Fixing a distribution $p \in \simplex$, we construct a set of $n$ witnesses in $\partial L(u,y)$ for each $y$ whose $p$-weighted sum is $\vec 0$.
%The nullspace of the matrix of these witnesses yields a flat, formally introduced below, of codimension at most $d$ intersecting our fixed distribution $p$ and the level set(s) $\gamma_r$ containing $p$.

@@

\bo{I like all this, but maybe it can be reworded to flow better with the above?}
Flats of a convex (expected) loss are complemented by its subgradients\footnote{We define the subgradient $\partial L(u,y) = \{g \in \reals^d : L(u',y) \geq L(u,y) +g^T (u' - u) \; \forall u' \in \reals^d  \}$.}.% of convex losses and their flats in order to understand convex elicitation complexity. 
While subgradients live in the prediction space and flats in probability space, the two can be connected considering a $p \in \simplex$.
In a loose sense, this $p$ is a witness of a flat of the level set, and the flat must be of certain codimension (dependent on prediction dimension) so that, for this witness $p$, we can extract a witness $W$ of subgradient sets proving that $\vec 0 \in \exploss{L}{u}{p}$; our desired optimality condition.\jessiet{This is probably TMI but it at least helped me to write it out and think of it this way (p as a witness of the flat, in a loose sense)}
The intuition for this result comes from the proof of~\citet[Theorem 16]{ramaswamy2016convex}, although indirect elicitation lets us simplify their proof that originally relied on calibration.
In fact, their main proof tool, the subspace of feasible dimensions (Definition~\ref{def:subspace-feas}), is a linear shift of a flat satisfying the requirements of Theorem~\ref{thm:cvx-flats}, which lets us deduce their result as Corollary~\ref{cor:fsd-bound} through a much shorter proof thanks to our reliance on elicitation.

@@

%Since Theorem~\ref{thm:consistent-implies-indir-elic} says that consistency implies indirect elicitation, we now apply results on convex elicitation complexity, like the result above, to understand the minimal dimension required to construct a consistent surrogate loss.
%This result suggests that elicitation complexity greater than $d$ implies no indirect elicitation of the property of interest (or target loss eliciting such a property), which in turn implies there is no $d$-dimensional consistent surrogate with respect to a given loss or property of interest.
\bo{I felt this remark was redundant now, so I removed it.}\jessie{Totally fine with that.}
